### 2022-10-08 workflow/notebooks/data-leakage-imputation.Rmd
updated w function

---
title: "Data leakage - does imputation before or after splitting matter?"
date: "`r Sys.Date()`"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.path = 'figures/')
```
```{r deps}
library(data.table)
library(here)
library(mikropml)
library(tidyverse)
```

## load data

```{r load_data}
allcause_full_dat <- fread(here('data','process','allcause_full_OTU.csv'))
```

TODO: also check attrib & idsa data sets

```{r count_num_nas}
count_num_nas <- function(dat) {
  temp <- colSums(is.na(dat)) 
  sum(temp)
}
```

```{r run_funcs}
count_num_nas(allcause_full_dat)
```

## is there any missing data?
```{r count_nas}
temp <- colSums(is.na(allcause_full_dat)) 
head(temp)
sum(temp)
```

## what's the distribution of missing data?

TODO histogram showing # of NA's in OTU columns instead of barplot (too many OTUs for barplot)
```{r plot_histogram}
missing.values <- allcause_full_dat %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 
missing.values %>%
  ggplot() +
    geom_histogram(aes(x=key, y=num.missing), stat = 'identity') +
    labs(x='variable', y="number of missing values", title='Number of missing values') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r check_works_with_simple_frame}
dat <- data.frame(x1 = c(NA, 5, 5, NA, 1, 2),
                   x2 = c(1, 2, 3, NA, 5, 6),
                   x3 = 1)
temp <- colSums(is.na(dat)) 
head(temp)
sum(temp)
missing.values <- dat %>%
    pivot_longer(everything(), names_to = "key", values_to = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 
missing.values %>%
  ggplot() +
    geom_histogram(aes(x=key), stat = 'count', binwidth=100) +
    labs(x='variable', y="number of missing values", title='Number of missing values') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## does the median value change before/after splitting?

### compute medians on entire dataset

### compute medians on just training set

```{r one-train-set, eval = FALSE}
training_inds <- get_partition_indices(allcause_full_dat %>% pull('allcause'),
        training_frac = 0.8,
        groups = NULL,
        group_partitions = NULL
      )
allcause_train_dat <- allcause_full_dat[training_inds, ]
```

### compare median differences -- histogram