### 2022-11-21 workflow/notebooks/ml-params.Rmd
updated plots

---
title: "ML parameters"
date: "`r Sys.Date()`"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.path = 'figures/')
```
```{r deps}
library(data.table)
library(here)
library(mikropml)
library(tidyverse)
```

```{r load_data}
results_dat <- data.table::fread(here('results', 
                                      'performance_results_aggregated.csv')
                                 ) %>% 
  rename(cv_metric_AUROC = cv_metric_AUC,
         AUROC = AUC,
         AUPRC = prAUC) %>% 
  mutate(metric = case_when(metric == 'AUC' ~ 'AUROC',
                            TRUE ~ metric)) %>% 
  pivot_longer(starts_with('cv_metric_'), 
               names_to = 'cv_metric', 
               values_to = 'cv_value') %>% 
  pivot_longer(c(AUROC, AUPRC, F1), 
               names_to = 'perf_metric_name',
               values_to = 'perf_metric_value') %>% 
  filter(!is.na(cv_value))
```

parameter choices to investigate:

- `trainfrac`: training set size as a fraction(`0.65`, `0.8`)
- `dataset`: all samples for all outcomes (`full`) vs intersection of outcomes not missing (`int`) 
- `method`: random forest (`rf`) vs L2 logistic regression (`glmnet`)
- `metric`: optimize model for `AUROC` vs `F1`
- taxonomic level: OTU vs genus. TODO: train models on genus level features

## training fraction

## dataset - full vs intersection

hold train frac, method, & metric constant

```{r plot_dataset}
results_dat %>% 
  filter(trainfrac == 0.8, method == 'glmnet', metric == 'AUROC') %>% 
  ggplot(aes(perf_metric_value, perf_metric_name, color = outcome)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap('dataset')

results_dat %>% 
  filter(trainfrac == 0.8, method == 'glmnet', metric == 'AUROC') %>% 
  ggplot(aes(perf_metric_value, dataset, color = outcome)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap('perf_metric_name')
```

## ML method - rf vs l2 logit
```{r plot_method}
results_dat %>% 
  filter(trainfrac == 0.8, metric == 'AUROC', dataset == 'full') %>% 
  ggplot(aes(perf_metric_value, perf_metric_name, color = outcome)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap('method')

results_dat %>% 
  filter(trainfrac == 0.8, metric == 'AUROC', dataset == 'full') %>% 
  ggplot(aes(perf_metric_value, method, color = outcome)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap('perf_metric_name')
```

## metric - AUROC vs F1 score
```{r plot_metric}
results_dat %>% 
  filter(trainfrac == 0.8, method == 'glmnet', dataset == 'full') %>% 
  ggplot(aes(perf_metric_value, perf_metric_name, color = outcome)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap('metric')

results_dat %>% 
  filter(trainfrac == 0.8, method == 'glmnet', dataset == 'full') %>% 
  ggplot(aes(perf_metric_value, metric, color = outcome)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap('perf_metric_name')
```