### 2023-02-13 workflow/notebooks/nmds-plots.Rmd
preliminary plot of nmds data

---
title: "nmds-plots"
author: "Megan Coden"
date: "2023-02-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("vegan")
# install.packages("readr")
```

```{r prepdata, include=FALSE}
library(vegan)
library(readr)
# source(here::here("code","plot","config-plots.R"))
# pc = read_tsv(here("data/mothur/cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
pc = read_tsv("/Users/megan/Desktop/schloss-lab/adverseCDI/data/mothur/cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
 #make community matrix - extract columns with abundance information
com = pc[,2:ncol(pc)] 
  #turn abundance data frame into a matrix
m_com = as.matrix(com)
set.seed(123)
minShift <- function(x) {x + abs(min(x))}
m_com <- apply(m_com, 2, minShift)
nmds = metaMDS(m_com, distance = "bray")
```

```{r plotnmds, echo=FALSE}
#extract NMDS scores (x and y coordinates)
data.scores = as.data.frame(scores(nmds))
#extract NMDS scores (x and y coordinates) for sites from newer versions of vegan package
data.scores = as.data.frame(scores(nmds)$sites)
#add columns to data frame 
data.scores$Sample = pc$Sample
data.scores$Time = pc$Time
data.scores$Type = pc$Type

head(data.scores)

library(ggplot2)

xx = ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) + 
    geom_point(data=data.scores)

 xx

ggsave("NMDS.svg")

```