### 2023-02-13 workflow/notebooks/nmds-plots.Rmd
fix nmds plots but still wrong data

---
title: "nmds-plots"
author: "Megan Coden"
date: "2023-02-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("vegan")
# install.packages("readr")
```

```{r prepdata, include=FALSE}
library(vegan)
library(readr)
library(tidyverse)
library(ggplot2)
# source(here::here("code","plot","config-plots.R"))
# pc = read_tsv(here("data", "mothur", "cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
pc = read_tsv("/Users/megan/Desktop/schloss-lab/adverseCDI/data/mothur/cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
case_data <- read.csv("/Users/megan/Desktop/schloss-lab/adverseCDI/data/process/cases_full_metadata.csv")
 #make community matrix - extract columns with abundance information
# com = pc[,2:ncol(pc)] 
  #turn abundance data frame into a matrix
# m_com = as.matrix(com)
# set.seed(123)
# minShift <- function(x) {x + abs(min(x))}
# m_com <- apply(m_com, 2, minShift)
# nmds = metaMDS(m_com, distance = "bray")
```

```{r split_by_outcomes, include=FALSE}
# inner join w sample id column from data/process/cases_full_metadata.csv
# plot will be one per definition, with severe as one color and not severe as other
names(pc)[names(pc) == "group"] <- "sample_id"
pc
combined_df <- pc %>% inner_join( case_data, 
           by=c('sample_id'))
idsa <- combined_df %>% select("axis1", "axis2", "idsa") %>% drop_na(idsa)
attrib <- combined_df %>% select("axis1", "axis2", "attrib") %>% drop_na(attrib)
allcause <- combined_df %>% select("axis1", "axis2", "allcause") %>% drop_na(allcause)

```

```{r plotnmds, echo=FALSE}
data.scores = as.data.frame(idsa)
idsa_plot <- ggplot(data.scores, aes(x = axis1, y = axis2)) + geom_point(aes(color = idsa))
data.scores = as.data.frame(attrib)
attrib_plot <- ggplot(data.scores, aes(x = axis1, y = axis2)) + geom_point(aes(color = attrib))
data.scores = as.data.frame(allcause)
allcause_plot <- ggplot(data.scores, aes(x = axis1, y = axis2)) + geom_point(aes(color = allcause))

idsa_plot
attrib_plot
allcause_plot

```