### 2023-02-13 workflow/notebooks/nmds-plots.Rmd
updated file reading to use here library

---
title: "nmds-plots"
author: "Megan Coden"
date: "2023-02-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("vegan")
# install.packages("readr")
library(vegan)
library(readr)
library(tidyverse)
library(ggplot2)
library(data.table)
library(here)
```

```{r prepdata, include=FALSE}
# source(here::here("code","plot","config-plots.R"))
# pc = read_tsv(here("data", "mothur", "cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
# pc = read_tsv("/Users/megan/Desktop/schloss-lab/adverseCDI/data/mothur/cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
pc <- data.table::fread(here('data', 'mothur',
                                      'cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes'))
#case_data <- read.csv("/Users/megan/Desktop/schloss-lab/adverseCDI/data/process/cases_full_metadata.csv")
case_data <- data.table::fread(here('data', 'process',
                                      'cases_full_metadata.csv'))
```

```{r split_by_outcomes, include=FALSE}
# inner join w sample id column from data/process/cases_full_metadata.csv
# plot will be one per definition, with severe as one color and not severe as other
names(pc)[names(pc) == "group"] <- "sample_id"
pc
combined_df <- pc %>% inner_join( case_data, 
           by=c('sample_id'))
idsa <- combined_df %>% select("axis1", "axis2", "idsa") %>% drop_na(idsa)
attrib <- combined_df %>% select("axis1", "axis2", "attrib") %>% drop_na(attrib)
allcause <- combined_df %>% select("axis1", "axis2", "allcause") %>% drop_na(allcause)

```

```{r plotnmds, echo=FALSE}
data.scores = as.data.frame(idsa)
idsa_plot <- ggplot(data.scores, aes(x = axis1, y = axis2)) + geom_point(aes(color = idsa))
data.scores = as.data.frame(attrib)
attrib_plot <- ggplot(data.scores, aes(x = axis1, y = axis2)) + geom_point(aes(color = attrib))
data.scores = as.data.frame(allcause)
allcause_plot <- ggplot(data.scores, aes(x = axis1, y = axis2)) + geom_point(aes(color = allcause))

idsa_plot
attrib_plot
allcause_plot

```