### 2023-02-20 workflow/notebooks/nmds-plots.Rmd

---
title: "nmds-plots"
author: "Megan Coden"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("vegan")
# install.packages("readr")
library(vegan)
library(readr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(data.table)
library(here)
```

```{r prepdata, include=FALSE}
# source(here::here("code","plot","config-plots.R"))
# pc = read_tsv(here("data", "mothur", "cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
# pc = read_tsv("/Users/megan/Desktop/schloss-lab/adverseCDI/data/mothur/cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes")
pc <- data.table::fread(here('data', 'mothur',
                                      'cdi.opti_mcc.braycurtis.0.03.lt.ave.nmds.axes'))
#case_data <- read.csv("/Users/megan/Desktop/schloss-lab/adverseCDI/data/process/cases_full_metadata.csv")
case_data <- data.table::fread(here('data', 'process',
                                      'cases_full_metadata.csv'))
```

```{r split_by_outcomes, include=FALSE}
# inner join w sample id column from data/process/cases_full_metadata.csv
# plot will be one per definition, with severe as one color and not severe as other
names(pc)[names(pc) == "group"] <- "sample_id"
pc
combined_df <- pc %>% inner_join( case_data, 
           by=c('sample_id'))
idsa <- combined_df %>% select("axis1", "axis2", "idsa") %>% drop_na(idsa) %>% rename(is_severe=idsa) %>% mutate(outcome="IDSA")
attrib <- combined_df %>% select("axis1", "axis2", "attrib") %>% drop_na(attrib) %>% rename(is_severe=attrib) %>% mutate(outcome="Attrib")
allcause <- combined_df %>% select("axis1", "axis2", "allcause") %>% drop_na(allcause) %>% rename(is_severe=allcause) %>% mutate(outcome="Allcause")
all_outcomes <- bind_rows(allcause, attrib, idsa)
```

```{r plotnmds, echo=FALSE}

data.scores = as.data.frame(all_outcomes)

plot <- data.scores %>% ggplot(aes(x=axis1, y=axis2, color = is_severe)) + geom_point(alpha = 0.7) + facet_wrap("outcome", nrow = 1) + theme_bw() + scale_color_manual(values=c(yes="#E41A1C", no="#377EB8"), name="is_severe") + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
plot
```