### 2023-02-23 workflow/notebooks/presentation.qmd

---
title: "Predicting _C. difficile_ infection severity from the taxonomic composition of the gut microbiome"
author: Kelly Sovacool
format: revealjs
title-slide-attributes:
    data-background-image: images/bacteria-rods-purple.jpeg
    data-background-size: contain
    data-background-opacity: "0.5"
embed-resources: true
---

::: {.notes}
title slide image source: https://www.mskcc.org/news/microorganisms-gut-can-affect-cancer-outcomes
:::

## Background

## _C. difficile_ infections (CDI) are on the rise

::: aside
https://www.cdc.gov/cdiff/clinicians/faq.html
:::

## Patient factors for predicting complicated CDI

- Electronic health record data used as features to predict whether disease-related complications occurred.
- Future application: inform clinicians on which CDI patients may be most at risk of severe CDI.