### 2023-03-17 workflow/notebooks/tax-search.Rmd
Plot ASVs too

---
title: "Taxonomic level search"
date: "`r Sys.Date()`"
output: 
  github_document:
    html_preview: false
  gfm:
    html-preview: false
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.path = 'figures/')
```
```{r deps}
library(data.table)
library(here)
library(mikropml)
library(schtools)
library(tidyverse)
library(yaml)
```

```{r load_data}
results_dat <- data.table::fread(here('results', 
                                      'performance_results_aggregated.csv')
                                 ) %>% 
  rename(train = cv_metric_AUC,
         test = AUC,
         AUPRC = prAUC) %>% 
  pivot_longer(c(train, test), names_to = 'set', values_to = 'AUROC') %>% 
  mutate(taxlevel = factor(taxlevel, levels = c("order", "family", "genus", "OTU", 'asv')))
```

## config

```{r config}
read_yaml(here('config', 'tax-search.yml'))
```

## model performance

```{r tax-search_perf}
results_dat %>% 
  ggplot(aes(x = taxlevel, y = AUROC, color = set)) +
  geom_hline(yintercept = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  facet_wrap(dataset~outcome)+#, nrow = 1) +
  theme_sovacool()
```