### 2023-04-25 workflow/notebooks/cost-benefit-analysis.Rmd
TODO: evaluate models with number needed to benefit

---
title: "Cost-Benefit Analysis"
date: "`r Sys.Date()`"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- use values from confusion matrix for one representative model on a test set.
  - number needed to screen
  - number needed to treat
  - number needed to benefit
- assumptions
  - cost of a non-severe case
  - cost of a severe case
    - days in icu, colectomy
  - cost of possible treatments:
    - abx: metronizadole, vancomycin, fidaxomycin
    - fmt, monoclonal antibodies

## evaluating ml models

[number needed to benefit](https://academic.oup.com/jamia/article-abstract/26/12/1655/5516459)

> In the simplest terms, prediction can be distilled into an NNS and action into a number needed to treat. When contextualized within this framework, the product of NNS and number needed to treat results in a number needed to benefit. The table outlines key variables in this framework that will alter the estimated number needed to benefit across different modeling and implementation scenarios.

## lit review

### [Gupta _et al.](https://journals.sagepub.com/doi/10.1177/17562848211018654)

- average cost of CDI case in the US: 
  - \$8k to \$30k (Nanwa _et al_)
  - \$21k (Zhang _et al_)
  - likely underestimates of true attributable costs.
- treatment   
  - IDSA recommends either vancomycin or fidaxomycin for 10 days
  - Metronizadole out of favor, not efficacious
  - Recommend FMT after multiple recurrences
  - monoclonal Ab now an fda-approved treatment

#### treatment costs

> More recently, Rajasingham et al. calculated the costs of the currently
> available therapies for CDI.
> The cost of oral metronidazole (10-day course) ranged from US$4.38 to
> US$13.14, intravenous metronidazole (14-day course) from US$19.56 to $58.68,
> vancomycin (10-day course) from US$7.04 to US$21.12, rifaximin (20-day course)
> from US$44.16 to US$132.48, and fidaxomicin (10day course), being the most
> expensive option, ranged from US$883.60 to US$2650.80. It is difficult to
> predict the exact cost of FMT due to the multiple variables involved, including
> source of stool and route of administration. In general, one course of FMT is
> estimated to cost between US$500 and US$2000.

### [fidaxomycin clinical trial](https://www.nejm.org/doi/full/10.1056/nejmoa0910812)

no significant difference in whether CDI was cured, but did have significant
difference in recurrence.

### [severe CDI treatment options](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3088840/)

**from 2008, now out of date**

see table for different treatment recommendations for mild-moderate CDI, 
IDSA severe CDI, and complicated CDI.

could compare cost-benefit analysis with IDSA severity.

### [idsa treatment guidelines comparison](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9780550/)

i may need to also consider "fulminant" cdi criteria

### [mich med treatment guidelines](https://www.med.umich.edu/1info/FHP/practiceguides/InptCDiff/C-Diff.pdf)

from 2019, has this been updated?