### 2023-05-31 paper/supplement.qmd
Create supplement

---
title: 'Supplement'
fontsize: 11pt
mainfont: Helvetica
format:
  pdf:
    keep-tex: false
    linenumbers: true
    doublespacing: true
    lof: false
crossref:
  fig-title: FIG S
---

::: {#fig-performance fig-cap="**Precision-recall curves.**"}
![](figures/prc_curves.png)
:::