### 2023-06-07 paper/supplement.qmd
Incorporate almost all of Pat's suggestions
Kept AUBPRC

<!-- TODO how get supplement figures labelled with S prefix
-->

::: {#fig-prc fig-cap="**Precision-recall curves.**
The original precision-recall curves for each model.
The horizontal line is the baseline precision, i.e. the proportion of severe
cases in the dataset for each severity definition.
Since each definition has a different baseline precision, the PRCs cannot be
compared directly without balancing the precision (see Figure 2).
"}
![](figures/prc_curves.png)
:::

::: {#fig-cdiff fig-cap="**_C. difficile_ relative abundance and feature importance.**
Of the 45 OTUs belonging to the _Peptostreptococcaceae_ family, only one
(OTU 25) had abundance values above the limit of detection.
**Left**: log~10~-transformed relative abundance of OTU 25 in the full datasets.
The dashed line is the limit of detection.
**Right**: Permutation feature importance as measured by AUROC for OTU 25.
The point is the mean difference in AUROC and the tails are the 75% confidence
interval.
The dotted line is a feature importance of zero, meaning the feature is not
important.
"}
![](figures/cdiff-otu.png)
:::

<!-- TODO supplementary figure with alpha and beta diversity & significance.
Samples were rarefied to 5,000 sequences per sample, repeated 1,000
times for alpha and beta diversity analysis.
 -->